Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data

### 2020-02-12 submission/cv_models.tex
fix typos after full read

In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current opensource pyTorch framework (version 1.4)~\cite{pytorch}.
This also includes a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.

@@

We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},% 
\footnote{That is, these test accuracies have been previously reported and made publicly-avaiable by others.  We take them as given, and we do not attempt to reproduce/verify them, since we do not permit ourselves any access to training/test data.}
as well as a basic linear regression line. 
%with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalziation (i.e., greater accuracy, lower error). 
While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).

@@

See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics on for the VGG, ResNet, and DenseNet series.
Similar results (not shown) are obtained for the Top5 accuracies.
Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.

@@

We are interested in how our four quality metrics depend on data set size.
To examine this, we look at results on ResNet versus ResNet-1K.
See Figure~\ref{fig:cv2-accuracy}, which plots and compares the Log $\alpha$-Norm metric 
%% MM: From Table 1 it is slightly worse than alpha-hat and we say that below %% --the best performing metric--
for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, which has been trained on a much smaller ImageNet-1K data set.
The Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
The ResNet series has strong correlation, with an RMSE of $0.66$, whereas the ResNet-1K series also shows good correlation, but has a much larger RMSE of $1.9$.
(Other metrics exhibit similar behavior.)
As expected, the higher quality data set shows a better fit, even with fewer data points.

@@

}    \subfigure[ ResNet-1K (overlaid) ]{
        \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
        \label{fig:resnet_alpha_overlaid_depth}
    }
    \caption{PL exponent ($\alpha$) versus layer id, for the least and the most accurate models in VGG (a), ResNet (b), and DenseNet (c) series. 
             (VGG is without BN; and note that the Y axes on each plot are different.)  
             Subfigure (d) displays the ResNet models (b), zoomed in to $\alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, from smallest to largest, to
             allow a more detailed analysis for the most strongly correlated layers.
             Notice that ResNet152 exhibits different and much more stable behavior of $\alpha$ across layers.
             This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
             In the text, this is interpreted in terms of \emph{Correlation Flow}.
            }
    \label{fig:3models-alpha-layers}
\end{figure}

@@

%Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
%VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.  
%The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
%%\ncyan{
\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing \emph{residual connections}.
This optimized approach allows for greater accurary with far fewer parameters (and GPU memory requirements), and
ResNet models of up to 1000 layers have been trained.\cite{resnet1000}.
%%}
We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
Contrast this with the DenseNet models, which contains many connections between every layer.
Our results (large $\alpha$, meaning they even a HT model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.

@@

We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package~\cite{distiller}.
We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models. 
The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$).  Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones. \emph{The oppopsite is observed.}
%\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).  }
%In both cases (Frobenius norm results not shown), we observe the opposite.
\charles{WHy is this confusing? Call me...}
Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%
\footnote{Here, we only include layer matrices or feature maps with $M\ge50$.}
On the other hand, the $\alpha$ values (in Figure~\ref{fig:resnet204Dalpha}) do not differ systematically between the baseline and fine-tuned models.
%\michael{Some comment about big jump in norm; but fix previous backwards question first.}--SEE BELOW
Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).

>>>>

In this section, we examine empirical quality metrics described in Section~\ref{sxn:methods} for several CV model architecture series.
This includes the VGG, ResNet, and DenseNet series of models, each of which consists of several pretrained DNN models, trained on the full ImageNet~\cite{imagenet} dataset, and each of which is distributed with the current open source pyTorch framework (version 1.4)~\cite{pytorch}.
This also includes a larger set of ResNet models, trained on the ImageNet-1K dataset~\cite{imagenet}, provided on the OSMR ``Sandbox for training convolutional networks for computer vision''~\cite{osmr}, which we call the ResNet-1K series.

@@

We have examined the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
To start, Figure~\ref{fig:vgg-metrics} considers the VGG series (in particular, the pretrained models VGG11, VGG13, VGG16, and VGG19, with and without BN), and it plots the four quality metrics versus the reported test accuracies~\cite{pytorch},% 
\footnote{That is, these test accuracies have been previously reported and made publicly-avaiable by others.  We take them as given, and we do not attempt to reproduce/verify them, since we do not permit ourselves any access to training/test data.}
as well as a basic linear regression line. 
%with the Root Mean Squared Error (RMSE) shown, \michael{XXX.  WHERE IS RMSE.  IN THAT TABLE?}
All four metrics correlate quite well with the reported Top1 accuracies, with smaller norms and smaller values of $\hat{\alpha}$ implying better generalization (i.e., greater accuracy, lower error). 
While all four metrics perform well, notice that the Log $\alpha$-Norm metric ($\log\Vert\mathbf{W}\Vert_{\alpha}^{\alpha}$) performs best (with an RMSE of $0.42$, see Table~\ref{table:cv-models}); and the Weighted Alpha metric ($\hat\alpha =\alpha\log\lambda_{max} $), which is an approximation to the Log $\alpha$-Norm metric~\cite{MM20_unpub_work}, performs second best (with an RMSE of $0.48$, see Table~\ref{table:cv-models}).

@@

See Table~\ref{table:cv-models} for a summary of results for Top1 accuracies for all four metrics for the VGG, ResNet, and DenseNet series.
Similar results (not shown) are obtained for the Top5 accuracies.
Overall, for the the ResNet, ResNet-1K, and DenseNet series, all metrics perform relatively well, the Log $\alpha$-Norm metric performs second best, and the Weighted Alpha metric performs best.
These model series are all well-trodden, and our results indicate that norm-based metrics and PL-based metrics can both distinguish among ``good-better-best'' models, with PL-based metrics performing somewhat (i.e., quantitatively) better.

@@

We are interested in how our four quality metrics depend on data set size.
To examine this, we look at results on ResNet versus ResNet-1K.
See Figure~\ref{fig:cv2-accuracy}, which plots and compares the Log $\alpha$-Norm metric 
%% MM: From Table 1 it is slightly worse than alpha-hat and we say that below %% --the best performing metric--
for the full ResNet model, trained on the full ImageNet dataset, against the ResNet-1K model, which has been trained on a much smaller ImageNet-1K data set.
The Log $\alpha$-Norm is much better than the Log Frobenius/Spectral norm metrics (although, as Table~\ref{table:cv-models} shows, it is actually slightly worse than the Weighted Alpha metric).
The ResNet series has strong correlation, with an RMSE of $0.66$, whereas the ResNet-1K series also shows good correlation, but has a much larger RMSE of $1.9$.
(Other metrics exhibit similar behavior.)
As expected, the higher quality data set shows a better fit, even with fewer data~points.

@@

}    \subfigure[ ResNet-1K (overlaid) ]{
        \includegraphics[width=4.0cm]{img/resnet_alpha_overlaid_depth.png} 
        \label{fig:resnet_alpha_overlaid_depth}
    }
    \caption{PL exponent ($\alpha$) versus layer id, for the least and the most accurate models in VGG (a), ResNet (b), and DenseNet (c) series. 
             (VGG is without BN; and note that the Y axes on each plot are different.)  
             Subfigure (d) displays the ResNet models (b), zoomed in to $\alpha\in[1,5]$, and with the layer ids overlaid on the X-axis, from smallest to largest, to
             allow a more detailed analysis of the most strongly correlated layers.
             Notice that ResNet152 exhibits different and much more stable behavior of $\alpha$ across layers.
             This contrasts with how both VGG models gradually worsen in deeper layers and how the DenseNet models are much more erratic.  
             In the text, this is interpreted in terms of \emph{Correlation Flow}.
            }
    \label{fig:3models-alpha-layers}
\end{figure}

@@

%Recall the differences---in particular, the number of residual connections---between the VGG, ResNet, and DenseNet architectures.
%VGG, while a fine model in many ways, was limited by needing massive FC layers at the end of the model, requiring significant memory and computational resources, and it's poor conditioning led to problems with vanishing gradients.  
%The empirical manifestations of this (on weight matrices) are that fitted $\alpha$ values get much worse for deeper layers.
%%\ncyan{
\paragraph{Comparison of VGG, ResNet, and DenseNet Architectures.}
We can interpret these observations by recalling the architectural differences between the VGG, ResNet, and DenseNet architectures,
and, in particular, the number of of residual connections. VGG resembles the  traditional convolutional architectures,
such as LeNet5, and consists of several [Conv2D-Maxpool-ReLu] blocks, followed by 3 large Fully Connected (FC) layers.
ResNet greatly improved on VGG by replacing the large FC layers, shrinking the Conv2D blocks, and introducing \emph{residual connections}.
This optimized approach allows for greater accurary with far fewer parameters (and GPU memory requirements), and
ResNet models of up to 1000 layers have been trained~\cite{resnet1000}.
%%}
We conjuecture that the effeciency and effectivness of ResNet is reflected in the smaller and more stable $\alpha\sim 2.0$, across nearly all layers, indicating that the inner layers are very well correlated and strongly optimized.
Contrast this with the DenseNet models, which contains many connections between every layer.
Our results (large $\alpha$, meaning they even a PL model is probably a poor fit) suggest that DenseNet has too many connections, diluting high quality interactions across layers, and leaving many layers very poorly optimized.

@@

We consider ResNet20, trained on CIFAR10, before and after applying the Group Regularization distillation technique, as implemented in the \texttt{distiller} package~\cite{distiller}.
We analyze the pretrained 4D\_regularized\_5Lremoved baseline and fine-tuned models. 
The reported baseline test accuracies (Top1$=91.45$ and Top5$=99.75$) are better than the reported fine-tuned test accuracies (Top1$=91.02$ and Top5$=99.67$).  Because the baseline accuracy is greater,  the previous results on ResNet (Table~\ref{table:cv-models} and Figure~\ref{fig:cv2-accuracy}) suggests that the baseline Spectral Norms should be \emph{smaller} on average than the fine-tuned ones. \emph{The oppopsite is observed.}
%\michael{Do we have something backwards here: where we observed that the average norm to increase with decreasing test error (as it ``should'' not), whereas the average PL exponent $\alpha$ decreases (as ``expected'' from HT-SR Theory).  }
%In both cases (Frobenius norm results not shown), we observe the opposite.
%%\charles{WHy is this confusing? Call me...}
Figure~\ref{fig:resnet204D5L} presents the Spectral Norm (here denoted $\log\lambda_{max}$) and PL exponent ($\alpha$) for each individual layer weight matrix $\mathbf{W}$.%
\footnote{Here, we only include layer matrices or feature maps with $M\ge50$.}
On the other hand, the $\alpha$ values (in Figure~\ref{fig:resnet204Dalpha}) do not differ systematically between the baseline and fine-tuned models.
%\michael{Some comment about big jump in norm; but fix previous backwards question first.}--SEE BELOW
Also (not shown), the average (unweighted) baseline $\alpha$ is \emph{smaller} than the fine-tuned average (as predicted by HT-SR Theory, the basis of $\hat{\alpha}$).

### 2020-02-12 submission/introduction.tex
fix typos after full read

A common problem in machine learning (ML) 
%and artificial intelligence (AI) 
is to evaluate the quality of a given model.
A popular way to accomplish this
%, in particular in academic environments, 
is to train a model and then evaluate its training/testing error.
There are many problems with this approach.
The training/testing curves give very limited insight into the overall properties of the model; 
they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling;
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability; 
and so on.
A less well-known problem, but one that is increasingly important, in particular in industrial-scale artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
Here, one may not have access to either the training data or the testing data.
Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it "as is,'' or to fine-tune and/or compress it and then use it.

@@

To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
Here is a summary of our main results.
\begin{itemize}
\item
\textbf{Norm-based metrics and well-trained models.}
Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
\item
\textbf{Norm-based metrics and poorly-trained models.}
Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.), exhibiting \emph{Scale Collapse} for these models. 
\item 
\textbf{PL-based metrics and model quality.}
PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
They are quantitatively better at discriminating ``good-better-best'' trends, and qualitatively better at distinguishing ``good-versus-bad'' models.
\item 
\textbf{PL-based metrics and model diagnostics.}
PL-based metrics can also be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
\end{itemize}

@@

All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository.%
\footnote{\url{https://github.com/CalculatedContent/kdd2020} \michael{TO BE ANONYMIZED.}}
See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.

>>>>

A common problem in machine learning (ML) 
%and artificial intelligence (AI) 
is to evaluate the quality of a given model.
A popular way to accomplish this
%, in particular in academic environments, 
is to train a model and then evaluate its training/testing error.
There are many problems with this approach.
The training/testing curves give very limited insight into the overall properties of the model; 
they do not take into account the (often large human and CPU/GPU) time for hyperparameter fiddling;
they typically do not correlate with other properties of interest such as robustness or fairness or interpretability; 
and so on.
A less well-known problem, but one that is increasingly important, in particular in industrial-scale artificial intelligence (AI), arises when the model \emph{user} is not the model \emph{developer}.
Here, one may not have access to either the training data or the testing data.
Instead, one may simply be given a model that has already been trained---a \emph{pretrained model}---and need to use it ``as is,'' or to fine-tune and/or compress it and then use it.

@@

To answer this question, we analyze hundreds of publicly-available pretrained state-of-the-art CV and NLP models. 
Here is a summary of our main results.
\begin{itemize}
\item
\textbf{Norm-based metrics and well-trained models.}
Norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
\item
\textbf{Norm-based metrics and poorly-trained models.}
Norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.), exhibiting \emph{Scale Collapse} for these models. 
\item 
\textbf{PL-based metrics and model quality.}
PL-based metrics do much better at predicting quality trends in pretrained CV/NLP models.  
They are quantitatively better at discriminating ``good-better-best'' trends, and qualitatively better at distinguishing ``good-versus-bad''~models.
\item 
\textbf{PL-based metrics and model diagnostics.}
PL-based metrics can also be used to characterize fine-scale model properties (including layer-wise \emph{Correlation Flow}) in well-trained and poorly-trained models, and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
\end{itemize}

@@

All of our computations were performed with the publicly-available \emph{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}.
To be fully reproducible, we only examine publically-available, pretrained models, and we also provide all Jupyter and Google Colab notebooks used in an accompanying github repository~\cite{kdd20_sub_repo}.
See Appendix~\ref{sxn:appendix} for details on how to reproduce all results.

### 2020-02-12 submission/methods.tex
fix typos after full read

For norm-based metrics, we use the average of the log norm, and to the appropriate power.
Informallly, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\Vert\cdot\Vert$ is a matrix norm.   
The log complexity,
\begin{equation}
\label{eqn:eqn:sum_log_norm}
\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert  ,
\end{equation}
 takes the form of an average Log Norm.
For the \emph{Frobenius norm metric} and \emph{Spectral norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly.% 
\footnote{When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.}

The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix
(and it approximates the average log $\alpha$-Shatten Norm metric),
\begin{equation}
\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
\end{equation}
where $L$ is the total nuumber of layer weight matrices.
The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.

Based on this, in this paper, we introduce and evaluate the $\alpha$-Shatten Norm metric.
Notice for the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum:

@@

One expects $\hat{\alpha}$ approximates the average log $\alpha$-Shatten Norm very well for $\alpha < 2$ and reasonably well for $\alpha\in[2,5]$~\cite{MM20_unpub_work}.

To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  
(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
The PL exponent $\alpha$ measures of the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
It is valid for $\lambda<\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
%
Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.%
\footnote{For small $\alpha$, this Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm, and may track this metric in some cases.}

>>>>

For norm-based metrics, we use the average of the log norm, and to the appropriate power.
Informallly, this amounts to assuming that the layer weight matrices are statistically independent, in which case we can estimate the model complexity $\mathcal{C}$, or test accuracy, with a standard Product Norm (which resembles a data dependent VC complexity),
\begin{equation}
\mathcal{C}\sim\Vert\mathbf{W}_{1}\Vert\times\Vert\mathbf{W}_{2}\Vert \times \cdots \times \Vert\mathbf{W}_{L}\Vert ,
\end{equation}
where $\Vert\cdot\Vert$ is a matrix norm.   
The log complexity,
\begin{equation}
\label{eqn:eqn:sum_log_norm}
\log\mathcal{C} \sim \log\Vert\mathbf{W}_{1}\Vert+\log\Vert\mathbf{W}_{2}\Vert + \cdots + \log\Vert\mathbf{W}_{L}\Vert  ,
\end{equation}
 takes the form of an average Log Norm.
For the \emph{Frobenius Norm metric} and \emph{Spectral Norm metric}, we can use Eqn.~(\ref{eqn:eqn:sum_log_norm}) directly.% 
\footnote{When taking $\log\Vert\mathbf{W}_{l}\Vert_{F}^{2}$, the $2$ comes down and out of the sum, and thus ignoring it only changes the metric by a constant factor.}

The \emph{Weighted Alpha metric} is an average of $\alpha_l$ over all layers $l \in \{1,\ldots,l\}$, weighted by the size, or scale, or each matrix,
%(and it approximates the average log $\alpha$-Shatten Norm metric),
\begin{equation}
\hat{\alpha} = \dfrac{1}{L}\sum_l \alpha_l\log\lambda_{max,l}\approx\langle\log\Vert\mathbf{X}\Vert_{\alpha}^{\alpha}\rangle    ,
\end{equation}
where $L$ is the total nuumber of layer weight matrices.
The Weighted Alpha metric was introduced previously~\cite{MM20_SDM}, where it was shown to correlate well with trends in reported test accuracies of pretrained DNNs, albeit on a limited set of models.

Based on this, in this paper, we introduce and evaluate the \emph{$\alpha$-Shatten Norm metric}.
Notice for the $\alpha$-Shatten Norm metric, however, $\alpha_l$ varies from layer to layer, and so in Eqn.~(\ref{eqn:sum_log_alpha_norm_alpha}) it can not be taken out of the sum:

@@

\noindent
For small $\alpha$, the Weighted Alpha metric approximates the Log $\alpha$-Shatten norm, as can be shown with a statistical mechanics and random matrix theory derivation \cite{MM20_unpub_work}; and the Weighted Alpha and $\alpha$-Shatten norm metrics often behave like an improved, weighted average Log Spectral Norm, and may track this metric in some~cases.

To avoid confusion, let us clarify the relationship between $\alpha$ and $\hat{\alpha}$.  
We fit the ESD of the correlation matrix $\mathbf{X}$ to a truncated PL, parameterized by 2 values: the PL exponent $\alpha$, and the maximum eigenvalue $\lambda_{max}$.  
(Technically, we also need the minimum eigenvalue $\lambda_{min}$, but this detail does not affect our analysis.)
The PL exponent $\alpha$ measures of the amount of correlation in a DNN layer weight matrix $\mathbf{W}$. 
It is valid for $\lambda<\lambda_{max}$, and it is scale-invariant, i.e., it does not depend on the normalization of $\mathbf{W}$ or $\mathbf{X}$.
The $\lambda_{max}$ is a measure of the size, or scale, of $\mathbf{W}$.
%
Multiplying each $\alpha$ by the corresponding $\log\lambda_{max}$ weighs ``bigger'' layers more, and averaging this product leads to a balanced, Weighted Alpha metric for the entire~DNN.
